% Essential Formatting

\documentclass[12pt]{article}
\usepackage{epsfig,amsmath,amsthm,amssymb,textcomp}
\usepackage[questions, answersheet]{../../urmathtest}[2001/05/12]

\def\ds{\displaystyle}

\newcommand{\ansbox}[1]
{\work{
  \pos\hfill \framebox[#1][l]{ANSWER:\rule[-.3in]{0in}{.7in}}
}{}}

\newcommand{\ansrectangle}
{\work{
  \pos\hfill \framebox[6in][l]{ANSWER:\rule[-.3in]{0in}{.7in}}
}{}}


% Beginning of the Document

\begin{document}
\examtitle{DATA MINING W4240}{HOMEWORK 4}{09/14/2010}

\begin{center}
	Professor: Frank Wood
\end{center}

% Problems Start Here % ----------------------------------------------------- %

{\bf Preliminary Instructions}

\begin{enumerate}
	\item Download the skeleton code for the assignment at \\  http://www.stat.columbia.edu/$\sim$fwood/w4240/Homework/index.html
	\item Unzip the downloaded material in an appropriate folder, something like w4240/hw4/
	\item Open MATLAB and navigate to the folder containing the downloaded material
\end{enumerate}

In this homework you need to implement variational Bayesian inference for the gaussian mixture model.  You will also need to implement the k-means algorithm.  The clustering obtained by k-means will then be used as an initialization for the variational algorithm.\\ 

\problem{25}{Implement a k-means clustering algorithm.  This should be done by filling out the two functions {\bf update\_cluster\_centers.m} and {\bf k\_means.m}.  The final return will be a list of cluster assignments, one for each data point.  After writing these functions you should be able to cluster $d$-dimensional data using the algorithm.  Since you will be using the output as an initialization you should program the algorithm in such a way that it returns $k$ clusters, i.e., that it does not collapse to fewer clusters than the input number $k$.  If you find that a cluster has no data points assigned to it during the algorithm, one potential solution is to assign that cluster mean to the value of a randomly chosen data point.  K-means is an iterative algorithm, but in this code you can hardcode the number of iterations to 10.\\ \\}


\problem{75}{Implement the variational Bayes inference algorithm for a gaussian mixture model and evaluate its performance on Fisher's Iris data.  To do this, you will need to implement the functions {\bf get\_r.m}, {\bf get\_other\_parameters.m}, and {\bf expected\_rand\_index.m}.  More explicit direction regarding the functionality of these programs can be found in the header of each file.  The naming of variables in these programs is designed to mirror the example given in your text book.  If you are confused about what certain variables are please refer there first before asking the TA.  You will also need to fill out values for $K$ and the parameters of the prior distributions in the file {\bf main.m}. You will notice in the main file that the output from the $k$-means clustering algorithm is used to inform the initialization of the variational Bayes algorithm.  Also, included in the HW material is a function which plots $d$-dimensional data by plotting the first two $PCA$ components.  This is just a trick to help in seeing the clusters in $d$-dimensional data when $d > 2$.\\ \\}

{\bf Submitting your HW}

You must complete this HW assignment on your own, you are not permitted to work with any one else on the completion of this task.  Your grade will reflect your ability to implement a working version of the procedure.  Submitted code must run on my machine in less than 3 minutes.  Grading will be automated and the submitted files will be run, therefore to submit the HW you will need to follow the following directions exactly.

\begin{enumerate}
	\item Send an email to w4240.fall2010.stat.columbia.edu@gmail.com
	\item {Attach your updated MATLAB files 
		\begin{enumerate}
			\item {\bf update\_cluster\_centers.m} 
			\item	 {\bf k\_means.m}
			\item  {\bf get\_r.m}
			\item {\bf get\_other\_parameters.m}
			\item  {\bf expected\_rand\_index.m}
			\item  {\bf main.m}
		\end{enumerate} It is imperative that the names be exactly as described here. There should be no folders attached, only raw .m files.  You may not attach other MATLAB code files. }
	\item The subject will be exactly your Columbia UNI followed by a colon followed by hw4.  For example, if the TA were submitting this homework the subject would read {\bf nsb2130:hw4}
	\item If you submit hw more than once, later files will overwrite earlier files.
\end{enumerate}

% Problems End Here % ------------------------------------------------------- %

\problemsdone
\end{document}

% End of the Document
